---
title: "01 - Reading in Data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
For my GOVT 110 course, I aim to examine public opinion on immigration from Mexico to the US based on data from the Latin American Public Project (LAPOP). 

In this step, I am reading in the data for the United States and Mexico from years 2004 - 2018 that can be found in their [database](http://datasets.americasbarometer.org/database/index.php?freeUser=true). I chose to read in both the factored and unfactored forms of the data so that both would be easily accessible for analysis. This means that many steps are repeated twice in the beginning, but the factored and unfactored versions will be joined together later in the process.

Below, you can find the suggested workflow in which you will be able to read in the desired datasets, format them into a tidy version, and explore the data as desired.

## Suggested workflow
1. Reading in Data
2. Categories and Questions
3. Adding Wave Column

## Libraries
Bringing in the same libraries used in [lapop-dashboard](https://github.com/vanderbilt-data-science/lapop-dashboard) repo that I worked on this past year. These packages handle reading in and clearning the data, and these will be the only ones you will need while following the suggest workflow.
```{r, message=FALSE}
library(labelled)
library(sjmisc)
library(tidyverse)
library(assertr)
library(haven)
library(janitor)
library(varhandle)
```

## Creating factored and unfactored dataframes
This function reads in and factors the given .dta file
```{r}
factor_country <- function(country){
  #Reads file for the country
  country <- read_dta(country)
  #Factors all the columns
  country <- as_factor(country)
  
  #Starts list x
  x <- 1
  
  #Changes name of column to add "factor"
  while(x <= length(country)){
      colnames(country)[x] <- paste(colnames(country)[x], "factor", sep = "_")
     x <- x + 1
  }
  country
}
```

# Reading in merged country files
I am using the **most recent** merged country file. Currently, there is no public file with 2016 and 2018 waves merged for Mexico, nor a public file with the 2018 data merged for the US.
```{r}
usa_factor <- factor_country("data/merged_usa_2006_2017.dta")
usa_unfactor <- read_dta("data/merged_usa_2006_2017.dta")
mexico_factor <- factor_country("data/merged_mexico_2004_2014.zip")
mexico_unfactor <- read_dta("data/merged_mexico_2004_2014.zip")
```

## Reading in 2016 file
Because there is no pre-existing merge that includes 2016 data, we are reading it in separately and then joining it later in the process.

If you are using a file that has this wave included, you can skip this.

### Creating factored data frame of 2016 questionnaire using `factor16_country`
This is the 2016 version of the `factor_country()` used above. The same process applies, but we have renamed the `uniq_id` column to `person_id` for clarification.
```{r}
factor16_country <- function(country){
  #Reads file for the country
  country <- read_dta(country)
  #Factors all the columns
  country <- as_factor(country) %>% 
    rename(person_id = uniq_id)
  
  #Starts list x
  x <- 1
  
  #Changes name of column to add "factor"
  while(x <= length(country)){
      colnames(country)[x] <- paste(colnames(country)[x], "factor", sep = "_")
     x <- x + 1
  }
  country
}
```


### Mexico 2016 File
Applying `factor16_country()` to the 2016 Mexico file.
```{r}
mexico_16_factor  <- factor16_country("data/2016_mexico.dta") %>% 
  #Checking for uniqueness
  assert(is_uniq, person_id_factor)

mexico_16_unfactor  <- read_dta("data/2016_mexico.dta") %>% 
  #Renaming ID column
  rename(person_id = uniq_id) %>% 
  #Checking for uniqueness
  assert(is_uniq, person_id) %>% 
  #Changing ID column to character class
  transform(person_id = as.character(person_id))
```

## Creating factored data frame of questionnarie using `factor18_country()`
This is similar to 2016 version above; the same process applies.
```{r}
factor18_country <- function(country){
  #Reads file for the country
  country <- read_dta(country)
  #Factors all the columns
  country <- as_factor(country) %>% 
    rename(person_id = uniq_id)
  
  #Starts list x
  x <- 1
  
  #Changes name of column to add "factor"
  while(x <= length(country)){
      colnames(country)[x] <- paste(colnames(country)[x], "factor", sep = "_")
     x <- x + 1
  }
  country
}
```

### USA 2018 file
```{r}
usa_18_factor <- factor18_country("data/2018_usa.dta") %>% 
  #Checking for uniqueness
  assert(is_uniq, person_id_factor)

usa_18_unfactor <- read_dta("data/2018_usa.dta") %>%
  #Renaming ID column
  rename(person_id = uniq_id) %>% 
  #Checking for uniquness
  assert(is_uniq, person_id) %>% 
  #Changning ID column to character class
  transform(person_id = as.character(person_id))
```

### Mexico 2018 file
```{r}
mexico_18_factor <- factor18_country("data/2018_mexico.dta") %>% 
  #Checking for uniqueness
  assert(is_uniq, person_id_factor)

mexico_18_unfactor <- read_dta("data/2018_mexico.dta") %>%
  #Renaming ID column
  rename(person_id = uniq_id) %>% 
  #Checking for uniquness
  assert(is_uniq, person_id) %>% 
  #Changning ID column to character class
  transform(person_id = as.character(person_id))
```

