---
title: "12 - Adding Wave Column"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
As in the LAPOP dashboard repository, the 2016 files did not include a wave column, so we had to add that manually. Here, I am adding wave column to the 2016 Mexico files.

The code comes from step 3 of the LAPOP dashboard repository, but is only applied to one country in this case. The merged US file accounted for 2016 in its wave_factor column.

## Function to adding wave column
### For unfactored datasets
```{r}
add_wave_unfactor_2016 <- function(country){
  #Adds new column, wave, with appropriate year
  country <- country %>%
    mutate(wave = 2016)
  return(country)
}
```

### For factored datasets
```{r}
add_wave_factor_2016 <- function(country){
  #Adds new column, wave, with appropriate year
  country <- country %>%
    mutate(wave_factor = 2016)
  return(country)
}
```

```{r}
#Bolivia
bolivia_16_factor <- add_wave_factor_2016(mexico_16_factor)
bolivia_16_unfactor <- add_wave_unfactor_2016(mexico_16_unfactor)
```

