---
title: "13 - Creating Unique ID"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
This follows the same code as step four in the LAPOP-dashboard repository. The 2016 and 2018 files arlready have a unique identifier, and we used this as a template for making the unique ID in the merged files.

First, there is a function written for the unfactored countries that is then applied to all unfactored merged countries.

Then, a similar function is written for the factored countries, but in this step we add "_factor" to the end of each column name to signifiy that it belongs to the factored dataset.

The unfactored and factored datasets will be joined later in the process.

## Function to make unique IDs for merged Mexico files
### Function to make unfactored unique ID for Mexico merged file
```{r}
making_uniqueID <- function(country){
  #Combines idnum and idnum_14 with "/"
  country <- mutate(country,
                    idnum = ifelse(idnum_14 == '',
                                   idnum,
                                   paste(idnum,idnum_14,sep = "_"))) %>%
    #Adds a underscore followed by a row number
    mutate(., idnum = paste(idnum, row_number(), sep = "_")) %>% 
    #Makes new column, person_id
    mutate(.,
           person_id = paste(
             paste(
               #Takes first 3 letters of country names and capitalizes it
               toupper(substr(pais, 1, 3)),
               #Selects last two digits of year
               substr(year, 3, 4), sep = ''),
             #Separates the country name from the year with underscore
             idnum, sep = "_"))
  
  #Converts person_id column to a character
  country <- transform(country, person_id = as.character(person_id)) %>% 
    assert(is_uniq, person_id)
  return(country)
}
```

### Function to make unique ID for factored data
The function below is the same as above, except `making_uniqueID` is now `making_uniqueID_factor` to indicate it needs to be used for factored data.
```{r}
making_uniqueID_factor <- function(country){
  #Combines idnum and idnum_14 with "/" and then makes unique id
  country <- mutate(country,
                    idnum_factor = ifelse(idnum_14_factor == '',
                                   idnum_factor,
                                   paste(idnum_factor,idnum_14_factor, sep = "_"))) %>%
    #Adds underscore with row number
    mutate(., idnum_factor = paste(idnum_factor, row_number(), sep = "_")) %>% 
    #Makes new column, uniq_id_factor
    mutate(.,
           person_id_factor = paste(
             paste(
               #Takes first 3 letters of country names and capitalizes it
               toupper(substr(pais_factor, 1, 3)),
               #Selects last two digits of year
               substr(year_factor, 3, 4), sep = ''),
             #Separates the country name from the year with underscore
             idnum_factor, sep = "_"))
  #Converts uniq_id column to a character
  country <- transform(country, person_id_factor = as.character(person_id_factor)) %>% 
    assert(is_uniq, person_id_factor)
  return(country)
}
```

## Applying `making_uniqueID` to merged Mexico file
### Unfactored, merged Mexico file
```{r}
mexico_unique_unfactor <- making_uniqueID(mexico_unfactor)
```

### Factored, merged Mexico file
```{r}
mexico_unique_factor <- making_uniqueID_factor(mexico_factor)
```

## Function to make unique IDs for merged Mexico files
### Factored, merged US
```{r}
making_uniqueID_factor <- function(country){
  #Combines idnum and idnum_14 with "/" and then makes unique id
  country <- mutate(country,
                    idnum_factor = ifelse(idnum_1416_factor == '',
                                   idnum_factor,
                                   paste(idnum_factor,idnum_1416_factor, sep = "_"))) %>%
    #Adds underscore with row number
    mutate(., idnum_factor = paste(idnum_factor, row_number(), sep = "_")) %>% 
    #Makes new column, uniq_id_factor
    mutate(.,
           person_id_factor = paste(
             paste(
               #Takes first 3 letters of country names and capitalizes it
               toupper(substr(pais_factor, 1, 3)),
               #Selects last two digits of year
               substr(year_factor, 3, 4), sep = ''),
             #Separates the country name from the year with underscore
             idnum_factor, sep = "_"))
  #Converts uniq_id column to a character
  country <- transform(country, person_id_factor = as.character(person_id_factor)) %>% 
    assert(is_uniq, person_id_factor)
  return(country)
}
```

### Uunfactored, merged US
```{r}
making_uniqueID <- function(country){
  #Combines idnum and idnum_14 with "/"
  country <- mutate(country,
                    idnum = ifelse(idnum_1416 == '',
                                   idnum,
                                   paste(idnum,idnum_1416,sep = "_"))) %>%
    #Adds a underscore followed by a row number
    mutate(., idnum = paste(idnum, row_number(), sep = "_")) %>% 
    #Makes new column, person_id
    mutate(.,
           person_id = paste(
             paste(
               #Takes first 3 letters of country names and capitalizes it
               toupper(substr(pais, 1, 3)),
               #Selects last two digits of year
               substr(year, 3, 4), sep = ''),
             #Separates the country name from the year with underscore
             idnum, sep = "_"))
  
  #Converts person_id column to a character
  country <- transform(country, person_id = as.character(person_id)) %>% 
    assert(is_uniq, person_id)
  return(country)
}
```

## Applying `making_uniqueID` to merged Mexico file
### Applying `making_uniqueID` to unfactored, merged US file
```{r}
usa_unique_unfactor <- making_uniqueID(usa_unfactor)
```

### Applying `making_uniqueID_factor` to US file
```{r}
usa_unique_factor <- making_uniqueID_factor(usa_factor)
```

