---
title: "15 - Data Exploration"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
I am now going to explore the clean versions of the Mexico and US data to look for patterns in the public opinion of immigration in both countries.

I am particularly interested in questions surrounding:
  + **who** are the people immigrating
  + why inidividuals are leaving their countries
  + how many individuals are actually migrating to/from each country
  + how citizens of each country view immigrants and immigration policy
  
## Reading through questions
In order to familiarize myself with the questions asked by LAPOP in these countries, I'll be using `tabyl()`. This will help give me insight in what qyestions and factors may be relevant.

## Making space for memory
```{r}
rm.all.but(keep = c("mexico_clean", "usa_clean"))
```

## Checking categories
```{r}
mexico_clean %>% 
  tabyl(category_en)
```

It looks like the possible pertinent categories are
  + Migration
  + Civic and Political Engagement
  + Climate Change/Disaster Risk Reduction Issues
  + Corruption
  + Crime, Insecurity, and Policy Preferences
  + Evaluation of Institutions, Political Systems and other Persons
  + Foreign Relations
  + Political and Religious Orientation
  + Political Tolerance
  + Political/Electoral Engagement
  + Socio-Demographic Characteristics
  + Support for Democracy and Democratic Institutions
  + The Economy

## Exploring questions by category and country
In order to narrow down the content, I'm going to look at the questions in each of these categories.

### Mexico
#### Migration
This would be definitely be necessary to analyze further. Asks about intentions of living or working abroad and remittances.
```{r}
# Overall questions in category
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```

#### Civic and Political Engagement
These only apply to political engagement in Mexico. I don't think that this would be super relevant to what I'm looking for.
```{r}
# Overall questions in category
mexico_clean %>% 
  filter(category_en == "Civic and Political Engagement") %>% 
  tabyl(question_es)
```


#### Climate Change/Disaster Risk Reduction Issues
I think here it could be interesting to see if people who are more at risk to natural disasters are more likely to immigrate.
```{r}
# Overall questions in category
mexico_clean %>% 
  filter(category_en == "Climate Change/Disaster Risk Reduction Issues") %>% 
  tabyl(question_es)

# Most relevant question
mexico_clean %>% 
  filter(question_es == "Probabilidad de muerte o daño en desastres naturales")
```


#### Corruption
This could be good for comparing how the citizens perceive their own goverments. I think that maybe higher perceived corruption could lead to an individual being more willing to emigrate.

I could clump all of the questions regarding bribes into one category and the other can be the perception of corruption.
```{r}
mexico_clean %>% 
  filter(category_en == "Corruption") %>% 
  tabyl(question_es)
```


#### Crime, Insecurity, and Policy Preferences
This should definitely be incorporated as well. In terms of comparison, the perception of retribution in the justice system could be compared between countries. Peception of danger within the neighborhood might be a bit too broad for comparison. However, it could be valuable to look at if the victims of crimes who have little faith in the judicial system are the ones leaving.
```{r}
mexico_clean %>% 
  filter(category_en == "Crime, Insecurity, and Policy Preferences") %>% 
  tabyl(question_es)
```


#### Evaluation of Institutions, Political Systems and other Persons
I feel like I could leave out questions about trust in the govermnet because it can be included in the perception of corruption. However, it may be worthwhile to keep the question about basic rights being protected and political freedom.
```{r}
mexico_clean %>% 
  filter(category_en == "Evaluation of Institutions, Political Systems and other Persons") %>% 
  tabyl(question_es)

# Keep these questions
mexico_clean %>% 
  filter(question_es == "Los derechos básicos están protegidos") 

mexico_clean %>% 
  filter(question_es == "Nivel actual de libertad para expresar opiniones políticas sin temor")
```


#### Foreign Relations
The question about the trustworthiness of the US government **may** or may not be useful. Probably not too relevant though.
```{r}
mexico_clean %>% 
  filter(category_en == "Foreign Relations") %>% 
  tabyl(question_es)
```


#### Political and Religious Orientation
I don't really think this is as important for Mexico, but I can see that this would be interesting for comparing how citizens in the US that identify with both ends of the political spectrum feel about migration into the US.
```{r}
mexico_clean %>% 
  filter(category_en == "Political and Religious Orientation") %>% 
  tabyl(question_es)
```


#### Political Tolerance
I don't think any of these have anything directly to do with immigration.
```{r}
mexico_clean %>% 
  filter(category_en == "Political Tolerance") %>% 
  tabyl(question_es)
```


#### Political/Electoral Engagement
Apparently there are no rows in this one?? Will have to come back to this.
```{r}
mexico_clean %>% 
  filter(category_en == "Political/Electoral Engagement") %>% 
  tabyl(question_es)
```


#### Socio-Demographic Characteristics
This will *definitely* be important for describing who are the people that are immigrating!!
```{r}
mexico_clean %>% 
  filter(category_en == "Socio-Demographic Characteristics") %>% 
  tabyl(question_es)
```


#### Support for Democracy and Democratic Institutions
This one isn't super relevant.
```{r}
mexico_clean %>% 
  filter(category_en == "Support for Democracy and Democratic Institutions") %>% 
  tabyl(question_es)
```


#### The Economy
This would be good for examining personal financial status.
```{r}
mexico_clean %>% 
  filter(category_en == "The Economy") %>% 
  tabyl(question_es)
```

### What to keep from Mexico
The categories to dig deeper into in Mexico are:
  + Migration
  + Climate Change/Disaster Risk Reduction Issues
  + Corruption
  + Crime, Insecurity, and Policy Preferences
  + Evaluation of Institutions, Political Systems and other Persons
  + Socio-Demographic Characteristics
  + Support for Democracy and Democratic Institutions
  + The Economy
  
### Removing unnecessary categories
I'm going to update mexico_clean so that it only contains the categories above.
```{r}
mexico_clean <- mexico_clean %>% 
  filter(category_en != "Civic and Politial Engagement") %>% 
  filter(category_en != "Foreign Relations") %>% 
  filter(category_en != "Political and Religious Orientation") %>% 
  filter(category_en != "Political Tolerance") %>% 
  filter(category_en != "Political/Electoral Engagement") %>% 
  filter(category_en != "Support for Democracy and Democratic Institutions")
```



### United States
#### Migration
This question is only asked once across all the data, but I wonder if this is because I only have the questions in Spanish...
```{r}
usa_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```

#### Civic and Political Engagement
I don't think any of these are particularly informative for this topic.
```{r}
usa_clean %>% 
  filter(category_en == "Civic and Political Engagement") %>% 
  tabyl(question_es)
```

#### Climate Change/Disaster Risk Reduction Issues
I'd say the last two mayyyy be helpful, but I think it's more important to examine that from the Mexican side.
```{r}
usa_clean %>% 
  filter(category_en == "Climate Change/Disaster Risk Reduction Issues") %>% 
  tabyl(question_es)
```

#### Corruption
I don't think this is necessary when examining from the US perspective.
```{r}
usa_clean %>% 
  filter(category_en == "Corruption") %>% 
  tabyl(question_es)
```

#### Crime, Insecurity, and Policy Preferences
These could be relevant for seeing if people who think immigrants are criminals have actually been victims of a crime.
```{r}
usa_clean %>% 
  filter(category_en == "Crime, Insecurity, and Policy Preferences") %>% 
  tabyl(question_es)

usa_clean %>% 
  filter(question_es == "Percepción de inseguridad en el barrio")

usa_clean %>% 
  filter(question_es == "Víctima de delincuencia en los últimos 12 meses")

usa_clean %>% 
  filter(question_es == "Víctima de delincuencia en los últimos 12 meses (Frecuencia)")
```

#### Evaluation of Institutions, Political Systems and other Persons
I don't think this counts much for immigration between Mex and US.
```{r}
usa_clean %>% 
  filter(category_en == "Evaluation of Institutions, Political Systems and other Persons") %>% 
  tabyl(question_es)
```

#### Foreign Relations
This has nothing to do with US-Mex relations.
```{r}
usa_clean %>% 
  filter(category_en == "Foreign Relations") %>% 
  tabyl(question_es)
```

#### Political and Religious Orientation
This would be important to see how side of the political spectrum feels about immigration
```{r}
usa_clean %>% 
  filter(category_en == "Political and Religious Orientation") %>% 
  tabyl(question_es)
```

#### Political Tolerance
Not relevant.
```{r}
usa_clean %>% 
  filter(category_en == "Political Tolerance") %>% 
  tabyl(question_es)
```

#### Political/Electoral Engagement
This is empty for some reason.
```{r}
usa_clean %>% 
  filter(category_en == "Political/Electoral Engagement") %>% 
  tabyl(question_es)
```

#### Socio-Demographic Characteristics
This would be important for seeing: if they are on welfare; age; ethnicity; and sex.
```{r}
usa_clean %>% 
  filter(category_en == "Socio-Demographic Characteristics") %>% 
  tabyl(question_es)

# Possible questions to focus on
usa_clean %>% 
  filter(question_es == "Beneficiario en el hogar del programa de transferencias de dinero condicionadas")

usa_clean %>% 
  filter(question_es == "Edad")

usa_clean %>% 
  filter(question_es == "Estado civil")

usa_clean %>% 
  filter(question_es == "Etnicidad")

usa_clean %>% 
  filter(question_es == "Ingreso mensual del hogar")

usa_clean %>% 
  filter(question_es == "Ocupación")

usa_clean %>% 
  filter(question_es == "Sexo")

usa_clean %>% 
  filter(question_es == "Situación laboral")
```

#### Support for Democracy and Democratic Institutions
Not necessary for US analysis.
```{r}
usa_clean %>% 
  filter(category_en == "Support for Democracy and Democratic Institutions") %>% 
  tabyl(question_es)
```

#### The Economy
It would be interesting to view their thoughts on the economic situation of the country.
```{r}
usa_clean %>% 
  filter(category_en == "The Economy") %>% 
  tabyl(question_es)

# Questions to keep
usa_clean %>% 
  filter(question_es == "Evaluación de la situación económica del país")

usa_clean %>% 
  filter(question_es == "Ingreso del hogar en los últimos dos años")

usa_clean %>% 
  filter(question_es == "Percepción de situación económica personal")
```

### What to keep from the US
  + Crime, Insecurity, and Policy Preferences
  + Political and Religious Orientation
  + Socio-Demographic Characteristics
  + The Economy
  
### Keeping relevant topics
First I'm going to make these into a list, and then I'm going to filter usa_clean based on if question_es is in the list
```{r}
cats_to_use <- c("Crime, Insecurity, and Policy Preferences", 
                 "Political and Religious Orientation", 
                 "Socio-Demographic Characteristics", 
                 "The Economy")

usa_clean <- usa_clean %>% 
  filter(category_en %in% cats_to_use)
```

