---
title: "15 - Data Exploration"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
I am now going to explore the clean versions of the Mexico and US data to look for patterns in the public opinion of immigration in both countries.

I am particularly interested in questions surrounding:
  + **who** are the people immigrating
  + why inidividuals are leaving their countries
  + how many individuals are actually migrating to/from each country
  + how citizens of each country view immigrants and immigration policy
  
## Reading through questions
In order to familiarize myself with the questions asked by LAPOP in these countries, I'll be using `tabyl()`. This will help give me insight in what qyestions and factors may be relevant.

## Making space for memory
```{r}
rm.all.but(keep = c("mexico_clean", "usa_clean"))
```

## Checking categories
```{r}
mexico_clean %>% 
  tabyl(category_en)
```

It looks like the possible pertinent categories are
  + Migration
  + Civic and Political Engagement
  + Climate Change/Disaster Risk Reduction Issues
  + Corruption
  + Crime, Insecurity, and Policy Preferences
  + Evaluation of Institutions, Political Systems and other Persons
  + Foreign Relations
  + Political and Religious Orientation
  + Political Tolerance
  + Political/Electoral Engagement
  + Socio-Demographic Characteristics
  + Support for Democracy and Democratic Institutions
  + The Economy

## Exploring questions by category and country
In order to narrow down the content, I'm going to look at the questions in each of these categories.

### Mexico
#### Migration
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```

#### Civic and Political Engagement
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Climate Change/Disaster Risk Reduction Issues
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Corruption
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Crime, Insecurity, and Policy Preferences
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Evaluation of Institutions, Political Systems and other Persons
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Foreign Relations
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Political and Religious Orientation
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Political Tolerance
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Political/Electoral Engagement
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Socio-Demographic Characteristics
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### Support for Democracy and Democratic Institutions
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```


#### The Economy
```{r}
mexico_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```

### United States
#### Migration
This question is only asked once across all the data, but I wonder if this is because I only have the questions in Spanish...
```{r}
usa_clean %>% 
  filter(category_en == "Migration") %>% 
  tabyl(question_es)
```

#### Civic and Political Engagement
#### Climate Change/Disaster Risk Reduction Issues
#### Corruption
#### Crime, Insecurity, and Policy Preferences
#### Evaluation of Institutions, Political Systems and other Persons
#### Foreign Relations
#### Political and Religious Orientation
#### Political Tolerance
#### Political/Electoral Engagement
#### Socio-Demographic Characteristics
#### Support for Democracy and Democratic Institutions
#### The Economy