---
title: "15 - Data Exploration"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

# Structure
I am now going to explore the clean versions of the Mexico and US data to look for patterns in the public opinion of immigration in both countries.

I am particularly interested in questions surrounding:
  + **who** are the people immigrating
  + why inidividuals are leaving their countries
  + how many individuals are actually migrating to/from each country
  + how citizens of each country view immigrants and immigration policy
  
## Reading through questions
In order to familiarize myself with the questions asked by LAPOP in these countries, I'll be using `tabyl()`. This will help give me insight in what qyestions and factors may be relevant.

## Making space for memory
```{r}
rm.all.but(keep = c("mexico_clean", "usa_clean"))
```

## Checking categories
```{r}
mexico_clean %>% 
  tabyl(category_en)
```

It looks like the possible pertinent categories are
  + Migration
  + Civic and Political Engagement
  + Climate Change/Disaster Risk Reduction Issues
  + Corruption
  + Crime, Insecurity, and Policy Preferences
  + Evaluation of Institutions, Political Systems and other Persons
  + Foreign Relations
  + Political and Religious Orientation
  + Political Tolerance
  + Political/Electoral Engagement
  + Socio-Demographic Characteristics
  + Support for Democracy and Democratic Institutions
  + The Economy
  
In order to narrow down the content, I'm going to look at the questions in each of these categories.